---
title: "Homework 02: Sexy Joe Biden"
author: Ellen Hsieh
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
```

### 1.
```{r}
biden = read.csv('/Users/ellenhsieh/Documents/UChicago/2019 Winter/Modeling/hw02/data/nes2008.csv')

qplot(biden$biden,
      geom="histogram",
      binwidth = 1, 
      main = "Distribution of Biden thermometer", 
      xlab = "Biden thermometer rating", 
      ylab = "Frequency"
      )
```

The distribution is skewed to the left. Although the respondents can choose any number to present their feeling thermometer from 0-100, most of them choose the number divisible by 5 or 10, not any random number like 37.

### 2.
```{r}
model1 = lm(biden ~ age, data = biden)
summary(model1)
cor(biden$biden, biden$age)
```

a. Is there a relationship between the predictor and the response?  
Yes, there is a relationship between age and biden feeling thermometer.

b. How strong is the relationship between the predictor and the response?  
The relationship is pretty weak. When the age increases by one, biden feeling thermometer increases only by 0.062.

c. Is the relationship between the predictor and the response positive or negative?  
It's positive.

d. Report the $R^2$ of the model. What percentage of the variation in biden does age alone explain? Is this a good or bad model?  
The $R^2$ for this model is 0.002, which indicates that this age model can only explain 0.002% of the variance in the biden feeling thermometer. 

e. What is the predicted biden associated with an age of 45? What are the associated 95% confidence intervals?  
```{r}
predict(model1, data.frame(age=45), level = 0.95, interval = 'confidence')
```
The 95% confidence interval for the predicted biden at age 45 is between 60.9 and 63.1. 

f. Plot the response and predictor. Draw the least squares regression line.
```{r}
ggplot(biden, aes(x = age, y = biden)) + 
  stat_smooth(method = "lm", col = "red") +
  geom_point(alpha = .25) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(title = "Linear relationship between Biden and Age",
        x = "Age",
        y = "Biden")
```

### 3.
```{r}
model2 = lm(biden ~ age + female + educ, data = biden)
summary(model2)
```

a. Is there a statistically significant relationship between the predictors and response?  
In this model, gender and education are strongly related to biden feeling thermometer. However, age is not statistically significant(p-value is 0.2 > 0.05)

b. What does the parameter for female suggest?  
This gender parameter suggests that women have marginally higher feeling towards Biden than men.

c. Report the $R^2$ of the model. What percentage of the variation in biden does age, gender, and education explain? Is this a better or worse model than the age-only model?  
This age-gender-education model explains 0.027% of the variance in biden feeling thermometer, which is higher than the previous one(0.002%). Thus, this model is better than the age-only model.  

d. Generate a plot comparing the predicted values and residuals, drawing separate smooth fit lines for each party ID type. Is there a problem with this model? If so, what?
```{r}
biden$fitted2 = fitted(model2)
biden$residual2 = residuals(model2)
biden$party = interaction(factor(biden$rep), factor(biden$dem))
levels(biden$party) = c('Independent', 'Republican', 'Democrat', 'both')
biden$party = droplevels(biden$party, exclude = 'both')

ggplot(biden, aes(fitted2, residual2, color = party)) + geom_point(aes(color = party)) + stat_smooth(aes(group = party), method = 'lm') +
labs(x = "Predicted Biden Feeling Thermometer",
     y = "Residuals")
```

The residuals for a good model should be scattered randomly, however, the residuals in this model are scattered in certain pattern. Also, the residuals for Democrat is higher than the other two.

### 4.
```{r}
model3 = lm(biden ~ age + female + educ + dem + rep, data = biden)
summary(model3)$coef
```

```{r}
biden$fitted3 = fitted(model3)
biden$residual3 = residuals(model3)
biden$party = interaction(factor(biden$rep), factor(biden$dem))
levels(biden$party) = c('none', 'rep', 'dem', 'both')
biden$party = droplevels(biden$party, exclude = 'both')
ggplot(biden, aes(fitted3, residual3, color = party)) + geom_point(aes(color = party)) + stat_smooth(aes(group = party), method = 'lm') +
labs(x = "Predicted Biden Feeling Thermometer",
     y = "Residuals")
```

The average of the predicted Biden feeling thermomoter among three parties are closer. However, none of the predicted values overlap acorss each other.

### 5.
```{r}
biden2 = biden %>%
  filter(dem + rep != 0)

model4 = lm(biden ~ factor(female) * factor(dem), data = biden2)
